Determining the alternatives for scalar implicature

### 2016-01-29 Paper/Paper.Rmd
minor edits, spelling, naming, before sending to MF

abstract: 
    Succesful communication regularly requires listeners to make pragmatic inferences - enrichments beyond the literal meaning of a speaker's utterance. For example, when interpreting a sentence such as "Alice ate some of the cookies," listeners routinely infer that Alice did not eat all of them. A Gricean account of this phenomena assumes the presence of alternatives (like "all of the cookies") with varying degrees of informativity, but it remains an open question precisely what these alternatives are. We use a computational model of pragmatic inference to test hypotheses about how well different sets of alternatives allow us to predict scalar implicature performance across a range of different scales. Our findings suggest that human comprehenders likely consider a much broader set of alternatives beyond those entailed by the initial description.

@@

@franke2014 used an RSA-style model to assess what alternatives a speaker would need to consider in order to produce the typicality/felicity ratings reported by @degen2015 and @vantiel2014. In order to do this, @franke2014's model assigned weights to a set of alternative numerical expressions. Surprisingly, along with weighting *one* highly (a conclusion that was suppored by the empirical work), the best-fitting model assigned substantial weight to *none* as an alternative. This finding was especially surprising considering the emphasis of standard theories on scalar items that stand in entailment relationships with one another (e.g. *one* entails *some* even if it is not classically considered to be part of the scale).

@@

We begin by presenting the computational framework we use throughout the paper. We next describe a series of experiments designed to measure both the literal semantics of a set of scalar items and comprehenders' pragmatic judgments for these same items. These experiments allow us to compare the effects of different alternative sets on our ability to model listeners' pragmatic judgements. To preview our results: we find that standard entailment alternatives do not allow us to fit participants' judgements, but that expanding the range of alternatives empirically (by asking participants to generate alternative messages) allows us to model listener judgements with high accuracy.

@@

We begin by giving a brief presentation of the basic RSA model. This model simulates the judgements of a pragmatic listener who wants to infer a speaker's intended meaning $m$ from her utterance $u$. For simplicity, we present a version of this model in which there is only full recursion: that is, the pragmatic listener reasons about a pragmatic speaker, who in turn reasons about a "literal listener." We assume throughout that this computation takes place in a signaling game [@lewis1969] with a fixed set of possible meanings $m \in M$ and a fixed possible set of utterances $u \in U$, with both known to both participants. Our goal in this study is to determine what utterances fall in $U$.

@@

This model of the pragmatic speaker is consistent with a speaker who chooses words to maximize the utility of an utterance in context [@frank2012], where utility is operationalized as the informativity of a particular utterance (surprisal) minus a cost:

@@

\noindent where $C(u)$ is the cost of a particular utterance, $-log(p_{L_0})$ represents the _surprisal_ for the literal listener (the information content of the utterance), and ($\alpha$ is a parameter in a standard choice rule. If $\alpha=0$, speakers choose randomly and as $\alpha \rightarrow \infty$, they greedily choose the highest probability alternative. In our simulations below, we treat $\alpha$ as a free parameter and fit it to the data.

To instantiate our signaling game with a tractable message set $M$, in our studies we adopt a food-review paradigm: we assume that speakers and listeners are trying to communicate the number of stars in an online restaurant review (where $m \in \{1, 2, 3, 4, 5\}$). We then use experiments to measure three components of the model. First, to measure literal semantics ${p_{L_0} (m \mid u)}$ (we ask experiment participants to judge whether a message is applicable to a particular meaning (Experiment 1). Second, to generate a set of plausible alternative messages in $U$, we elicit alternatives directly (Experiment 2). Lastly, to obtain human $L_1$ pragmatic judgments, we ask participants to interpret a speaker's utterances (Experiment 3).

@@

Experiment 1 was conducted to approximate literal listener semantic distributions $p_{L_0}(m \mid u)$ for five pairs of scalar items taken from @vantiel2014. We include three conditions in Experiment 1, corresponding to the sets of alternatives within a scale that participants were presented with: two alternatives ("entailment"), four alternatives, and five alternatives. The two-alternative entailment condition makes a test of the hypothesis that the two  members of the classic Horn scale [@horn1972] are the only alternatives necessary to predict the strength of listeners' pragmatic inference. The four- and five-alternative conditions then add successively more alternatives to test whether including a larger number of alternatives will increase model fit.\footnote{Note that alternatives in the four- and five-alternatives conditions were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.} A secondary goal of Experiments 1b,c tests whether the total set of alternatives queried during literal semantic elicitation effects compatibility judgments. (If it does we should see differences in compatibility judgments between experiments.) We address this possibiliy in the results section.

@@

Figure \ref{fig:stimuli_exp1}, left, shows the experimental setup. Participants were presented with a target scalar item and a star rating (1--5 stars) and asked to judge the compatibility of the scalar item and star rating. Compatibility was assessed through a binary "yes/no" response to a question of the form "Do you think that the person thought the food was \____?" where a target scalar was presented in the blank. Each participant saw all scalar item and star rating combinations for their particular condition, in a random order.

@@

```{r exp1Plots, fig.env = "figure*", fig.pos = "t", fig.width=6.5, fig.height=5, fig.align='center', set.cap.width=T, num.cols.cap=2, fig.cap = "Literal listener judgements from Experiments 1a,b,c. Proportion of participants indicating acceptability is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Each line-type shows a different experiment, and colors indicate the different items in the scale (with the experiments including different numbers of items)."}

@@

# for plotting
speakers$study <- ifelse(speakers$expt == "e8", "1a",
                                 ifelse(speakers$expt == "e10", "1b", "1c"))
ggplot(speakers, 
       aes(x = stars, y = speaker.p, col = study, lty = study)) + 
  geom_line() + 
  # facet_grid(.~ scale) + 
  facet_wrap(~wordPlots) +
  guides(colour = guide_legend(direction = "horizontal", nrow = 1, 
                               title = "Scalar Item"), 
         linetype = guide_legend(direction = "horizontal", nrow = 1, 
                                 title = "Experiment")) + 
  xlab("star rating") +
  ylab("proportion compatible ('yes' responses)") + 
  # ggtitle("Scalar item literal semantic compatibility\nwith star ratings") +
  # scale_colour_solarized() + 
  theme_bw() +
  theme(strip.text.x = element_text(size = 6),
        legend.position = "bottom",
        legend.margin = unit(.02,"in"),
        legend.key = element_blank(), 
        legend.key.size = unit(.1, "in"),
        legend.title = element_text(size = 10),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 6))
```

Figure \ref{fig:exp2Plots} plots literal listener $p_{L0}(m|u)$ scalar item distributions for the three conditions. Each row plots a unique scalar family with items ordered horizontally by valence. Several trends are visible. First, in each scale, the alternatives spanned the star scale, such that there were alternatives that were highly compatible with both the lowest and highest numbers of stars. Second, there was clear variability between scalar families. For example, compatibility judgments for the top items in the *memorable / unforgettable* scale were more similar than those for *good / excellent* or *liked / loved.* Finally, there was substantial consistency in ratings for items that were repeated across experiments, suggesting that this paradigm elicited stable judgments from participants. 
```{r literalListener_ME_models, echo = FALSE, warning = FALSE, message = FALSE}
# fix contrasts so we're comparing most disparate experimetns (e8 and e12)
contrasts(allLiteralListener$exp) <- cbind("e10" = c(1, 0, 0), "e8" = c(0, 0, 1))

# glmer WITHOUT experiment and with random effect for subjects
glm1 <- glmer(judgment ~ scale + stars + (1 | word) + (1 | workerid),
              family=  "binomial", data = allLiteralListener)
# glmer WITH experiment and with random effect for subjects
glm2 <- glmer(judgment ~ exp + scale +  stars + (1 | word) +  (1 | workerid),
              family = "binomial", data = allLiteralListener)
sumGlm2 <- summary(glm2)
anovaResults <- anova(glm1, glm2)
```
To test our secondary hypothesis, that the different sets of scalar items queried might result in differences in compatibility judgments between experiments, we ran a mixed effects model. We regressed compatibility judgments on scale, number of stars and experiment (1a,b,c), with subject and word level random effects, which was the maximal structure that converged. Results indicate no significant differences between Experiment 1a and 1c $b =$`r round(sumGlm2$coefficients[3], 2)`, $Z =$ `r round(sumGlm2$coefficients[19], 2)`, $p =$ `r round(sumGlm2$coefficients[27], 2)` or between 1b and 1c $b =$ `r round(sumGlm2$coefficients[2], 2)`, $Z =$ `r round(sumGlm2$coefficients[18], 2)` $p =$ `r round(sumGlm2$coefficients[26], 2)` and the addition of our experiment predictor did not significantly improve model fit when compared to a model without the experiment variable using ANOVA $\chi^2($ `r anovaResults$"Chi Df"[2]` $) =$ `r round(anovaResults$Chisq[2], 2)`, $p =$ `r round(anovaResults$"Pr(>Chisq)"[2], 2)`.

@@

Participants were presented a target scalar term from our original entailment set embedded in a sentence such as, "In a recent restaurant review someone said they thought they the food was \____," with a target scalar presented in the blank. Participants were then asked to generate plausible alternatives by responding to the question, "If they'd felt differently about the food, what other words could they have used instead of \____?" They were prompted to generate three unique alternatives.

@@

```{r exp2_altsPlot_likedLoved, fig.pos = "t", fig.align='center', set.cap.width=T, fig.height=2, num.cols.cap=2, fig.cap = "Combined counts for participant generated alternatives for 'liked \ loved' scale. Participants were shown a target scalar (either liked or loved) and asked to give three alternatives."}

@@

Experiment 3 was conducted to measure pragmatic judgments $p_{L_1}(m \mid u)$ for five pairs of scalar items taken from @vantiel2014. We include two conditions in Experiment 3. In the two-alternative condition, participants made judgments for items included in the entailment scales. In the four-alternative condition, participants made judgments for the entialment items and also the top two alternatives elicited for each scale in Experiment 2. Running two pragmatic judgment conditions allowed us to rule out the potential effects of having a larger set of alternatives during the pragmatic judgment elicitation and also provided two sets of human judgments to compare with model predictions. \footnote{Note that alternatives in the four-alternative condition were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; both pragmatic listener experiments are grouped together for simplicity in reporting.} A secondary goal of Experiment 3b tests whether the total set of alternatives queried in Experiments 3a,b effects pragmatic judgments. (If it does we should see differences in pragmatic judgments between experiments.) We address this possibiliy in the results section.

@@

Figure \ref{fig:exp2Plots} plots pragmatic listener judgments distributions for "weak" / "strong" scalar pairs (ie *good*/*excellent*). We only include the original entailment pairs in this figure for simplicity. Several trends are visible. First, in each scale participants generated implicatures and were less likely to assign high star-ratings to weaker scalar terms, despite literal semantic compatibility. Second, the "size" of the difference between judgments between "strong" and "weak" scalar items varied by scale, confirming results found by Tiel (2014).

To test our secondary hypothesis, that the different sets of scalar items queried might result in differences in pragmatic judgments between experiments, we ran a mixed effects model. We regressed pragmatic judgements on scale and experiment (3a,b) with subject and word level random effects, which was the maximal structure that converged. Results indicate there were no significant differences between Experiment 3a and 3b for $b =$ `r round(sumGlm2$coefficients[2], 2)`, $t($ `r round(sumGlm2$coefficients[14], 0)` $) =$ `r round(sumGlm2$coefficients[20], 2)`, $p =$ `r round(sumGlm2$coefficients[26], 2)` and the addition of our experiment predictor did not significantly improve model fit when compared to a model without the experiment variable using ANOVA $\chi^2($ `r anovaResults$"Chi Df"[2]` $) =$ `r round(anovaResults$Chisq[2], 2)`, $p =$ `r round(anovaResults$"Pr(>Chisq)"[2], 2)`.

@@

By varying the type of scale representations available to our Bayesian model we investigated the effects of alternatives on scalar implicature. Model fit with human judgments was significantly improved by the inclusion of alternatives beyond the typical "strong" and "weak" scalar items. The "Two alternative" model contained only entailment items, which, under classic accounts, should be sufficient to generate implicature. Model performance with this limited alternative set was poor when compared to human judgments from Experiment 3a,b (row 1, Table \ref{xtable}). Building off findings in Franke (2014), in which the best-fitting model assigned substantial weight to *none* as an alternative, we included a "generic" negative alternative that was only compatible with 1-star. Model fit for this modified "Two alternative with generic negative" alternative set did improve, however overall correlation was still low (row 2, Table \ref{xtable}). Model fit jumped substantially by adding the top-two alternatives for each scale (row 3, Table \ref{xtable}). Finally, our "Five alternative" model included a neutral alternative along with the original entailment pair and top two (row 4, Table \ref{xtable}).

While model improvement seemed to be related to the number of alternatives present, we did not see substantial differences between our four- and five-alternative models. One possibility is that alternatives are differentially salient in context. Our current framework makes the simplifying assumption that all alternatives are equally salient. However, it may well be the case that while both *all* and *most* are *present* during pragmatic enrichment of *some*, *all* is more salient and therefore should carry greater weight in our model. Findings from Experiment 2, in which participants generated plausible alternatives, supports this intuition, as some alternatives appeared more often than others. This distinction between binary *presence* and graded *salience* is an important one that deserves further investigation.

It is also likely that the precise set of alternatives present during implicature are domain dependent. Our current empirical paradigm elicited literal semantics, pragmatic judgments and plausible alternatives all within the restricted domain of food reviews. Participant data may have differed greatly if we had instead asked them to respond in a movie review paradigm, which would have been acceptable for a number of our stimuli (ie *memorable*/*unforgettable*, *liked*/*loved*, etc.).

@@

Thanks to NSF BCS XYZ. Thanks to Michael Franke, Judith Degen, and Noah Goodman.

>>>>

abstract: 
    Successful communication regularly requires listeners to make pragmatic inferences - enrichments beyond the literal meaning of a speaker's utterance. For example, when interpreting a sentence such as "Alice ate some of the cookies," listeners routinely infer that Alice did not eat all of them. A Gricean account of this phenomena assumes the presence of alternatives (like "all of the cookies") with varying degrees of informativity, but it remains an open question precisely what these alternatives are. We use a computational model of pragmatic inference to test hypotheses about how well different sets of alternatives allow us to predict scalar implicature performance across a range of different scales. Our findings suggest that human comprehenders likely consider a much broader set of alternatives beyond those entailed by the initial description.

@@

@franke2014 used an RSA-style model to assess what alternatives a speaker would need to consider in order to produce the typicality/felicity ratings reported by @degen2015 and @vantiel2014. In order to do this, @franke2014's model assigned weights to a set of alternative numerical expressions. Surprisingly, along with weighting *one* highly (a conclusion that was supported by the empirical work), the best-fitting model assigned substantial weight to *none* as an alternative. This finding was especially surprising considering the emphasis of standard theories on scalar items that stand in entailment relationships with one another (e.g. *one* entails *some* even if it is not classically considered to be part of the scale).

@@

We begin by presenting the computational framework we use throughout the paper. We next describe a series of experiments designed to measure both the literal semantics of a set of scalar items and comprehenders' pragmatic judgments for these same items. These experiments allow us to compare the effects of different alternative sets on our ability to model listeners' pragmatic judgments. To preview our results: we find that standard entailment alternatives do not allow us to fit participants' judgments, but that expanding the range of alternatives empirically (by asking participants to generate alternative messages) allows us to model listener judgments with high accuracy.

@@

We begin by giving a brief presentation of the basic RSA model. This model simulates the judgments of a pragmatic listener who wants to infer a speaker's intended meaning $m$ from her utterance $u$. For simplicity, we present a version of this model in which there is only full recursion: that is, the pragmatic listener reasons about a pragmatic speaker, who in turn reasons about a "literal listener." We assume throughout that this computation takes place in a signaling game [@lewis1969] with a fixed set of possible meanings $m \in M$ and a fixed possible set of utterances $u \in U$, with both known to both participants. Our goal in this study is to determine what utterances fall in $U$.

@@

This model of the pragmatic speaker (denoted $S_1$) is consistent with a speaker who chooses words to maximize the utility of an utterance in context [@frank2012], where utility is operationalized as the informativity of a particular utterance (surprisal) minus a cost:

@@

\noindent where $C(u)$ is the cost of a particular utterance, $-log(p_{L_0})$ represents the _surprisal_ for the literal listener (the information content of the utterance), and $\alpha$ is a parameter in a standard choice rule. If $\alpha=0$, speakers choose randomly and as $\alpha \rightarrow \infty$, they greedily choose the highest probability alternative. In our simulations below, we treat $\alpha$ as a free parameter and fit it to the data.

To instantiate our signaling game with a tractable message set $M$, in our studies we adopt a food-review paradigm: we assume that speakers and listeners are trying to communicate the number of stars in an online restaurant review (where $m \in \{1, 2, 3, 4, 5\}$). We then use experiments to measure three components of the model. First, to measure literal semantics ${p_{L_0} (m \mid u)}$ (we ask experiment participants to judge whether a message is compatible with a particular meaning (Experiment 1). Second, to generate a set of plausible alternative messages in $U$, we elicit alternatives directly (Experiment 2). Lastly, to obtain human $L_1$ pragmatic judgments, we ask participants to interpret a speaker's utterances (Experiment 3).

@@

Experiment 1 was conducted to approximate literal listener semantic distributions $p_{L_0}(m \mid u)$ for five pairs of scalar items taken from @vantiel2014. We include three conditions in Experiment 1, corresponding to the sets of alternatives within a scale that participants were presented with: two-alternatives ("entailment"), four-alternatives, and five-alternatives. The two-alternatives entailment condition makes a test of the hypothesis that the two  members of the classic Horn scale [@horn1972] are the only alternatives necessary to predict the strength of listeners' pragmatic inference. The four- and five-alternatives conditions then add successively more alternatives to test whether including a larger number of alternatives will increase model fit.\footnote{Note that alternatives in the four- and five-alternatives conditions were chosen on the basis of Experiment 2, which was run chronologically after the two-alternative condition; all literal listener experiments are grouped together for simplicity in reporting.} A secondary goal of Experiments 1a,b,c tests whether the set of alternatives queried during literal semantic elicitation impacts compatibility judgments. (If it does we should see differences in compatibility judgments for shared items between experiments.) We address this possibility in the results section.

@@

Figure \ref{fig:stimuli_exp1}, left, shows the experimental setup. Participants were presented with a target scalar item and a star rating (1--5 stars) and asked to judge the compatibility of the scalar item and star rating. Compatibility was assessed through a binary "yes/no" response to a question of the form, "Do you think that the person thought the food was \____?" where a target scalar was presented in the blank. Each participant saw all scalar item and star rating combinations for their particular condition, in a random order.

@@

```{r exp1Plots, fig.env = "figure*", fig.pos = "t", fig.width=6.5, fig.height=5, fig.align='center', set.cap.width=T, num.cols.cap=2, fig.cap = "Literal listener judgments from Experiments 1a,b,c. Proportion of participants indicating compatibility (answering 'yes') is shown on the vertical axis, with the horizontal axis showing number of stars on which the utterance was judged. Rows are grouped by scale and items within rows are ordered by valence. Colors indicate the specific experiment (1a,b,c) with experiments including different numbers of items."}

@@

# for plotting
speakers$study <- ifelse(speakers$expt == "e8", "1a",
                                 ifelse(speakers$expt == "e10", "1b", "1c"))
ggplot(speakers, 
       aes(x = stars, y = speaker.p, col = study)) + 
  geom_line() + 
  # facet_grid(.~ scale) + 
  facet_wrap(~wordPlots) +
  guides(colour = guide_legend(direction = "horizontal", nrow = 1, 
                               title = "Scalar Item"), 
         linetype = guide_legend(direction = "horizontal", nrow = 1, 
                                 title = "Experiment")) + 
  xlab("star rating") +
  ylab("proportion compatible ('yes' responses)") + 
  # ggtitle("Scalar item literal semantic compatibility\nwith star ratings") +
  # scale_colour_solarized() + 
  theme_bw() +
  theme(strip.text.x = element_text(size = 6),
        legend.position = "bottom",
        legend.margin = unit(.02,"in"),
        legend.key = element_blank(), 
        legend.key.size = unit(.1, "in"),
        legend.title = element_text(size = 10),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 6))
```

Figure \ref{fig:exp2Plots} plots literal listener $p_{L0}(m|u)$ scalar item distributions for the three conditions. Each row shows a unique scalar family with items ordered horizontally by valence. Several trends are visible. First, in each scale, the alternatives spanned the star scale, such that there were alternatives that were highly compatible with both the lowest and highest numbers of stars. Second, there was clear variability between scalar families. For example, compatibility judgments for the top items in the *memorable / unforgettable* scale were more similar than those for *good / excellent* or *liked / loved.* Finally, there was substantial consistency in ratings for items that were repeated across experiments, suggesting that this paradigm elicited stable judgments from participants. 
```{r literalListener_ME_models, echo = FALSE, warning = FALSE, message = FALSE}
# fix contrasts so we're comparing most disparate experimetns (e8 and e12)
contrasts(allLiteralListener$exp) <- cbind("e10" = c(1, 0, 0), "e8" = c(0, 0, 1))

# glmer WITHOUT experiment and with random effect for subjects
glm1 <- glmer(judgment ~ scale + stars + (1 | word) + (1 | workerid),
              family=  "binomial", data = allLiteralListener)
# glmer WITH experiment and with random effect for subjects
glm2 <- glmer(judgment ~ exp + scale +  stars + (1 | word) +  (1 | workerid),
              family = "binomial", data = allLiteralListener)
sumGlm2 <- summary(glm2)
anovaResults <- anova(glm1, glm2)
```
To test our secondary hypothesis, that the different sets of scalar items queried might result in differences in compatibility judgments between experiments, we ran a mixed effects model. We regressed compatibility judgments on scale, number of stars and experiment (1a,b,c), with subject and word level random effects, which was the maximal structure that converged. Results indicate no significant differences between Experiment 1a and 1c, $b =$`r round(sumGlm2$coefficients[3], 2)`, $Z =$ `r round(sumGlm2$coefficients[19], 2)`, $p =$ `r round(sumGlm2$coefficients[27], 2)` or between 1b and 1c, $b =$ `r round(sumGlm2$coefficients[2], 2)`, $Z =$ `r round(sumGlm2$coefficients[18], 2)` $p =$ `r round(sumGlm2$coefficients[26], 2)` and the addition of our experiment predictor did not significantly improve model fit when compared to a model without the experiment variable using ANOVA $\chi^2($ `r anovaResults$"Chi Df"[2]` $) =$ `r round(anovaResults$Chisq[2], 2)`, $p =$ `r round(anovaResults$"Pr(>Chisq)"[2], 2)`.

@@

Participants were presented a target scalar item from our original entailment set (see Figure \ref{fig:allScalesTable}) embedded in a sentence such as, "In a recent restaurant review someone said they thought they the food was \____," with a target scalar presented in the blank. Participants were then asked to generate plausible alternatives by responding to the question, "If they'd felt differently about the food, what other words could they have used instead of \____?" They were prompted to generate three unique alternatives.

@@

```{r exp2_altsPlot_likedLoved, fig.pos = "t", fig.align='center', set.cap.width=T, fig.height=2, num.cols.cap=2, fig.cap = "Combined counts for participant generated alternatives for 'liked \ loved' scale from Experiment 2. Participants were shown a target scalar (either liked or loved) and asked to give three alternatives."}

@@

Experiment 3 was conducted to measure pragmatic judgments $p_{L_1}(m \mid u)$ for five pairs of scalar items taken from @vantiel2014. We include two conditions in Experiment 3. In the two-alternatives condition, participants made judgments for items included in the entailment scales. In the four-alternatives condition, participants made judgments for the entailment items and also the top two alternatives elicited for each scale in Experiment 2. Running two pragmatic judgment conditions allowed us to rule out the potential effects of having a larger set of alternatives during the pragmatic judgment elicitation and also provided two sets of human judgments to compare with model predictions\footnote{Note that alternatives in the four-alternatives condition were chosen on the basis of Experiment 2, which was run chronologically after the two-alternatives condition; both pragmatic listener experiments are grouped together for simplicity in reporting.}. A secondary goal of Experiment 3b tests whether the set of alternatives queried in Experiments 3a,b impacts pragmatic judgments. (If it does we should see differences in pragmatic judgments between experiments.) We address this possibility in the results section.

@@

Figure \ref{fig:exp2Plots} plots pragmatic listener judgments distributions for "weak" / "strong" scalar pairs (e.g. *good*/*excellent*). We only include the original entailment pairs in this figure for simplicity. Several trends are visible. First, in each scale participants generated implicatures and were less likely to assign high star-ratings to weaker scalar terms, despite literal semantic compatibility. Second, the "size" of the difference between judgments between "strong" and "weak" scalar items varied by scale, consistent with results from Tiel (2014).

To test our secondary hypothesis, that the different sets of scalar items queried might result in differences in pragmatic judgments between experiments, we ran a mixed effects model. We regressed pragmatic judgments on scale and experiment (3a,b) with subject and word level random effects, which was the maximal structure that converged. Results indicate there were no significant differences between Experiment 3a and 3b, $b =$ `r round(sumGlm2$coefficients[2], 2)`, $t($ `r round(sumGlm2$coefficients[14], 0)` $) =$ `r round(sumGlm2$coefficients[20], 2)`, $p =$ `r round(sumGlm2$coefficients[26], 2)` and the addition of our experiment predictor did not significantly improve model fit when compared to a model without the experiment variable using ANOVA $\chi^2($ `r anovaResults$"Chi Df"[2]` $) =$ `r round(anovaResults$Chisq[2], 2)`, $p =$ `r round(anovaResults$"Pr(>Chisq)"[2], 2)`.

@@

By varying the type of scale representations available to our Bayesian model we investigated the effects of alternatives on scalar implicature. Model fit with human judgments was significantly improved by the inclusion of alternatives beyond the typical "strong" and "weak" scalar items. The two-alternatives model contained only entailment items, which, under classic accounts, should be sufficient to generate implicature. Model performance with this limited alternative set was poor when compared to human judgments from Experiment 3a,b (row 1, Table \ref{xtable}). Building off findings in Franke (2014), in which the best-fitting model assigned substantial weight to *none* as an alternative, we included a "generic" negative alternative that was only compatible with 1-star. Model fit for this modified entailment alternative set did improve, however overall correlation was still low (row 2, Table \ref{xtable}). Model fit jumped substantially by adding the top-two alternatives for each scale (row 3, Table \ref{xtable}). Finally, our five-alternatives model included a neutral alternative along with the original entailment pair and top two (row 4, Table \ref{xtable}).

While model improvement seemed to be related to the number of alternatives present, we did not see substantial differences between our four- and five-alternative models. One possibility is that alternatives are differentially salient in context. Our current framework makes the simplifying assumption that all alternatives are equally salient. However, it may well be the case that while both *all* and *most* are *present* during pragmatic enrichment of *some*, *all* is more salient and therefore should carry greater weight in our model. Findings from Experiment 2, in which participants generated plausible alternatives, support this intuition, as some alternatives appeared more often than others. This distinction between binary *presence* and graded *salience* is an important one that deserves further investigation.

It is also likely that the precise set of alternatives present during implicature are domain dependent. Our current empirical paradigm elicited literal semantics, pragmatic judgments and plausible alternatives all within the restricted domain of restaurant reviews. Participant data may have differed greatly if we had instead asked them to respond in a movie review paradigm, which would have been acceptable for a number of our stimuli (e.g. *memorable*/*unforgettable*, *liked*/*loved*, etc.).

@@

Thanks to NSF BCS. Thanks to Michael Franke, Judith Degen, and Noah Goodman.
